Embedded Scalars, Preferred Readings \& Prosody: {A}n Experimental Revisit

### 2015-10-17 text/paper/reviews_2/response_letter.md

# Response to editor and reviewer comments

The main changes from the previous version are:

1. careful reformulation of conclusions to be drawn from our data in line with reviewer 2's suggestions; concretely:

	a. we stress (even more than before) that grammaticalism qua "core theory" is compatible with our data, while traditionalism is not;

	b. still, we maintain that grammaticalism qua "core theory" is *trivially* compatible with our data, because it would not be inconsistent with any possible outcome of our experiment;

	c. the only auxiliary assumption for grammaticalism that is directly applicable to our experiment (selecting reading preferences in terms of logical strength alone) is not compatible with our data.

2. shortening of section 3 to remove unnecessary detail in predictions of "core theories"
3. addition of a few elements of emphasis in introduction, Section 4 and reflection section 6, in response to reviewer 2's comments showing that the previous version may not have made the motivation of our design sufficiently clear

@@

## Editor comments

>I would like to ask two things of you: 
>
>(1) have a good think about the balanced conclusion you draw (the data are hard for both sides of the theoretical divide --- reviewer 2: the data clearly favours grammaticalism) and address the reviewer's worry as good as you can (s/he can't be the only one interpreting your paper like that);

We agree that our previous version has not been ideally clear on this. It is not our intention to polarize by misleading rhetorics. We have tried to make very clear in the current version that our main conclusions from our data are (as already stated above):

a. we stress (even more than before) that grammaticalism qua "core theory" is compatible with our data, while traditionalism is not;

b. still, we maintain that grammaticalism qua "core theory" is *trivially* compatible with our data, because it would not be inconsistent with any possible outcome of our experiment;

c. the only auxiliary assumption for grammaticalism that is directly applicable to our experiment (selecting reading preferences in terms of logical strength alone) is not compatible with our data.

@@

>(2) rethink the value of the fine-grained description of the landscape of theoretical options (section 3.1/3.2).

Although we had thought that it would be good for the readers to see the full space of possible predictions about preference patterns that relevant "core theories" + "auxiliary hypotheses" give, we agree that it is not strictly necessary to introduce all of these distinctions. We have therefore removed those chunks of Section 3 which are not strictly relevant for later discussions (including Table 2, which summarized the predictions). Some particular "auxiliary assumptions" are taken up in the discussion section 6, but only those that are relevant for the overall discussion, like Magri's appeal to obligatory exhaustification, which is an example for how grammaticalism could predict our data.

### 2015-10-17 text/paper/revision_02/paper.tex
micha final edits version 2

Summing up, three experimental studies have presented diverging pieces
of evidence. % \citet{GeurtsPouscoulous2009:Embedded-Implic} did not
% observe any truth value-judgments that would indicate local readings
% in a picture verification task and thus conclude that these readings
% are unavailable. \citet{CliftonDube2010:Embedded-Implic} and
% \citet{ChemlaSpector2010:Experimental-Ev} employed different designs
% intended to be sensitive enough to reveal local readings. Both studies
% found putative evidence in support of local readings.
Differences in results might be due to differences in experimental design, in particular due to
the type of elicited judgement and the possibility of conflating pictorial effects. Also, we
should reflect on the effects of ``silent prosody.''

@@

\paragraph{Upshot.} This leaves us with the following desiderata. (i) In order to test the
availability and preferences of different readings of \as- and \es-sentences, we need a way to
unambiguously map responses, ideally categorical, to specific readings. (ii) We should try to
avoid issues of ``silent prosody'' and actively explore the role of prosodic markedness. (iii)
We would like to obtain information about the relative preferences of the different readings,
ideally by comparison to some other, ``base-line'' condition. Finally (iv) we want to minimize
potential picture typicality effects.

@@

\paragraph{General Idea.} To deal with these desiderata, we used an \emph{incremental
  verification task} (\acro{ivt}), which is a modified version of picture verification
\citep[see][]{Conroy2008}. The general idea is that subjects are requested to judge sentence
material with respect to pictures that do not necessarily contain all the information relevant
for judging a certain reading true or false. In that case, participants can demand that more
information be revealed. Participants are instructed to make a truth-value judgment as soon as
they are able to. When they do, the trial ends.

@@

\noindent If we compare (\ref{bsp:observed-preferences}) with
Table~\ref{tab:predictions}, we quickly see that these findings are
not predicted by either traditionalism or grammaticalism under the
auxiliary assumptions that we considered so far. We address each
position separately in the following.

Traditionalism easily explains the absence of global readings. Weak
traditionalism, introduced in Section~\ref{sec:traditionalism}, for
instance, would hold that the extra assumptions needed to derive
global readings (e.g., a speaker competence assumption) are not
available or only allow for epistemically weak implicatures that would
show as responses indicative of literal readings in our task. In
support of this view, traditionalists could argue that the task's
rather high processing demands lead to a suppression of global
readings, in line with the findings of
\citet{NeysDe-NeysSchaeken2007:When-People-Are} that working memory
load negatively affects the number of scalar implicature responses in
sentence verification tasks. The finding of local readings in
\as-sentences with contrastive stress is easily explained by adopting
the prosodic markedness hypothesis. What traditionalism does not
predict is the high number of answers indicative of local readings in
the \es-conditions. Resorting to the prosodic markedness hypothesis
does not help, because local readings were observed for accented as
well as for unaccented \es-sentences.%  It would also remain unclear why
% prosodic cues should have such differential effects on \as- and
% \es-sentences.

Grammaticalism, on the other hand, fails to predict the pattern in
(\ref{bsp:observed-preferences}) entirely if it equates preference for
readings with logical strength. Grammaticalism predicts that literal
readings of \as-sentences are the least preferred, contrary to
(\ref{bsp:observed-preferences}). Again, we need to bear the caveat of
Section~\ref{sec:grammaticalism} in mind that we are only evaluating
grammaticalism with respect to a particular disambiguation criterion
here. The verdict, however, seems to be clear. If
(\ref{bsp:observed-preferences}) shows the actual preference orders
over readings, then meaning disambiguation in terms of logical
strength is generally on the wrong track, because no account that
merely looks at the logical strengths of readings can plausibly
predict the preference pattern in (\ref{bsp:observed-preferences}) for
both \as- and \es-sentences at the same time.

In sum, the present findings cannot be explained in terms of the
versions of traditionalism and grammaticalism that we have considered
so far, i.e., with the particular auxiliary assumptions adopted in
Section~\ref{sec:theories-predictions}. But we should ask what
plausible amendements or alternative auxiliary assumptions would
enable either core theory to accommodate the data \emph{ex post}. This
is what we do in the next section, where we also reflect critically on
our design and our interpretation of the data.

@@

Section~\ref{sec:theories-predictions} introduced
traditionalism and grammaticalism as two competing core theories,
whose concrete empirical predictions depend on additional auxiliary
assumptions. We chose, in part for practical reasons, particular
auxiliary assumptions, but we should also ask more generally: on the
supposition that our design and the offered interpretation of our data
are sound, what would it take to accommodate the observed data under
the core theories?

\paragraph{Grammaticalism.} Consider grammaticalism first. Our data
contradicts the notion that preferences follow logical strength. But,
as noted in Section~\ref{sec:theories-predictions}, grammaticalism
could be supplemented by a conceptually different disambiguation
criterion. Selecting readings in terms of how well they answer the
contextually given question under discussion, as suggested by
\citet{Fox2007:Free-Choice-and,GualminiHulsey2008:The-Question-An}, is
an option. But it is quite unclear what the question under discussion
should be that guided judgements in our particular task. To make this
approach yield \emph{ex ante} predictions, we would need to determine
before the data is available which implicit QUDs our particular task
should give rise to, so as to derive a preference ordering over
readings. This may well be possible, but we do not presently see how.

@@

On conceptual grounds we are very much in favor of a disambiguation
criterion in terms of economy considerations, but whether the
particular account sketched here is empirically successful in other
cases as well, must remain to be seen. Two things are worth
emphasizing here nonetheless. Firstly, the sketched Magri-style
account is a plausible alternative \emph{post hoc} explanation. It was
not, at least to our knowledge, a salient possibility before seeing
our data set: other specifications of $Alt_1$ and $Alt_2$ could have
been made equally plausible \emph{ex ante}. Secondly, what our data
refutes is that preferences for readings follow logical
strength. This, therefore, does not jeopardize grammaticalism as a
core theory, but calls for theory-internal revision of its
disambiguation criterion. Taken together, we argue that the data
presented here points to a dire need of refining grammaticalism,
perhaps along the lines suggested here, but ideally alongside more
experimental data.

@@

\begin{figure}[t]
	\centering
	\subfloat[][Step 2  (\lit true)]{ 
		\fbox{\includegraphics[width=3.5cm]{pictures/PostExp_2.pdf}}
	    \label{fig:exseqES1}
	}
	\subfloat[][Step 3]{
		\fbox{\includegraphics[width=3.5cm]{pictures/PostExp_3.pdf}}
	    \label{fig:exseqES2}
	}
	\subfloat[][Step 4 (\acro{prag} false)]{ 
		\fbox{\includegraphics[width=3.5cm]{pictures/PostExp_4.pdf}}
	    \label{fig:exseqES3}
	}
	\caption{Example sequence for critical trial in
          post-experiment.}
	\label{fig:Trial-PostExp}
\end{figure}

@@

Backed up by our analyses, we concluded that both traditionalism and
grammaticalism do not predict our data under the set of auxiliary
assumptions that we provided them with in
Section~\ref{sec:theories-predictions}. Traditionalism fails to
account for high numbers of responses indicating local readings even
for prosodically unmarked \es-sentences. Grammaticalism, insofar as it
builds on disambiguation by logical strength, fails to predict the
absence of answers indicative of global readings and the relative
abundance of answers indicative of literal readings.

The clearest upshot of theoretical importance of this study is that
strength-based meaning selection does not seem to be a good
disambiguation criterion for grammaticalism. This point is also
relevant in general, because disambiguation based on logical strength
has been suggested in other domains as well
\citep[e.g.][]{DalrympleKanazawa1998:Reciprocal-Expr,Winter2001:Plural-Predicat,CobrerosEgre2012:Tolerant-Classi}.
We tentatively suggested an alternative selection principle for
grammaticalism in terms of ``parsing economy,'' but it remains to be
seen whether this is empirically successful in other domains as well.

@@

We have argued that auditory presentation of sentence material is crucial if
prosodic information is hypothesized to influence
availability and preferences of readings. We also argued that
theoretical positions must be formulated in such a way that they make
testable predictions about reading preferences where different
potential readings are allowed. We should then try to obtain
information about actual preference patterns by explicit comparison
with constructions whose preference structures are well-studied, ideally
with some model of the data generating process for optimal data
analysis.

>>>>

Summing up, three experimental studies have presented diverging pieces
of evidence. % \citet{GeurtsPouscoulous2009:Embedded-Implic} did not
% observe any truth value-judgments that would indicate local readings
% in a picture verification task and thus conclude that these readings
% are unavailable. \citet{CliftonDube2010:Embedded-Implic} and
% \citet{ChemlaSpector2010:Experimental-Ev} employed different designs
% intended to be sensitive enough to reveal local readings. Both studies
% found putative evidence in support of local readings.
Differences in results might be due to differences in experimental design, in particular due to
the type of elicited judgement and the possibility of conflating pictorial effects. Also, we
should consider potential effects of ``silent prosody.''

@@

\paragraph{Upshot.} This leaves us with the following desiderata. (i) In order to test the
availability and preferences of different readings of \as- and \es-sentences, we need a way to
unambiguously map responses, ideally categorical, to specific readings. (ii) We want to
minimize possible pictorial effects. (iii) We would like to obtain information about the
relative preferences of the different readings, ideally by comparison to some other,
``base-line'' condition. (iv) We should try to avoid issues of ``silent prosody'' and actively
explore the role of prosodic markedness.

@@

\paragraph{General motivation.} To deal with desiderata (i) and (ii), we used an
\emph{incremental verification task} (\acro{ivt}), which is a modified version of picture
verification \citep[see][]{Conroy2008}. The general idea is that subjects are requested to
judge sentence material with respect to pictures that do not necessarily contain all the
information relevant for judging a certain reading true or false. In that case, participants
can demand that more information be revealed. Participants are instructed to make a truth-value
judgment as soon as they are able to. When they do, the trial ends. Motivated by desideratum
(iii), we include ``preference-related control'' conditions. Sentence material was presented
auditorily and we manipulated prosodic realization to cater for desideratum (iv).

@@

\noindent Let's consider whether traditionalism or grammaticalism are comfortable with these
preference relations.

Traditionalism easily explains the absence of global readings by holding that the extra
assumptions needed to derive global readings (e.g., a speaker competence assumption) are not
available or only allow for epistemically weak implicatures that would show as responses
indicative of literal readings in our task. In support of this view, traditionalists could
argue that the task's rather high processing demands lead to a suppression of global readings,
in line with the findings of \citet{NeysDe-NeysSchaeken2007:When-People-Are} that working
memory load negatively affects the number of scalar implicature responses in sentence
verification tasks. The finding of local readings in \as-sentences with contrastive stress is
easily explained by adopting the prosodic markedness hypothesis. What traditionalism does not
predict is the high number of answers indicative of local readings in the
\es-conditions. Resorting to the prosodic markedness hypothesis does not help, because local
readings were observed for accented as well as for unaccented \es-sentences.

Grammaticalism, on the other hand, fails to predict the pattern in
(\ref{bsp:observed-preferences}) entirely if it equates preference for readings with logical
strength. Strength-based disambiguation predicts that literal readings of \as-sentences are the
least preferred, contrary to (\ref{bsp:observed-preferences}). Again, we need to bear the
caveat of Section~\ref{sec:grammaticalism} in mind that we are only evaluating grammaticalism
with respect to a particular disambiguation criterion here. The verdict, however, seems to be
clear. If (\ref{bsp:observed-preferences}) shows the actual preference orders over readings,
then meaning disambiguation in terms of logical strength is generally on the wrong track,
because no account that merely looks at the logical strengths of readings can plausibly predict
the preference pattern in (\ref{bsp:observed-preferences}) for both \as- and \es-sentences at
the same time.

In sum, the present findings appear to challenge traditionalism as a ``core theory'', and to
challenge grammaticalism under the ``auxiliary assumption'' that reading preferences mirror
logical strength. Given this, we should ask what plausible amendements or alternative auxiliary
assumptions would enable either position to accommodate the data \emph{ex post}. This is what
we do in the next section, where we also reflect critically on our design and our
interpretation of the data.

@@

Section~\ref{sec:theories-predictions} introduced traditionalism and grammaticalism as two
competing core theories, whose concrete empirical predictions depend on additional
assumptions. Here, we should finally ask more generally: on the supposition that our design and
the offered interpretation of our data are sound, what would it take to accommodate the
observed data under the core theories?

\paragraph{Grammaticalism.} Consider grammaticalism first. Our data contradicts the notion that
preferences follow logical strength. But, as noted in Section~\ref{sec:theories-predictions},
grammaticalism could be supplemented by a conceptually different disambiguation
criterion.\footnote{See also the discussion by \citet{ChemlaSpector2010:Experimental-Ev} on the
  prospects of pushing strength-based disambiguation in the light of their data, which they
  also take to imply contradicting evidence on reading preferences.} Selecting readings in
terms of how well they answer the contextually given question under discussion, as suggested by
\citet{Fox2007:Free-Choice-and,GualminiHulsey2008:The-Question-An}, is an option. But it is
quite unclear what the question under discussion should be that guided judgements in our
particular task.

@@

On conceptual grounds we are very much in favor of a disambiguation criterion in terms of
economy considerations, but whether the particular account sketched here is empirically
successful in other cases as well, must remain to be seen. Two things are worth emphasizing
here nonetheless. Firstly, the sketched account is a plausible \emph{post hoc} explanation. It
was not, at least to our knowledge, a salient possibility before seeing our data set: other
specifications of $Alt_1$ and $Alt_2$ could have been made equally plausible \emph{ex
  ante}. Secondly, what our data refutes is that preferences for readings follow logical
strength. This, therefore, does not jeopardize grammaticalism as a core theory, but merely
calls for theory-internal revision of its disambiguation criterion, ideally alongside more
experimental data.

@@

\begin{figure}[t]
	\centering
	\subfloat[][Step 2  (\lit true)]{ 
		\fbox{\includegraphics[width=3.5cm]{pictures/PostExp_2.pdf}}
	    \label{fig:exseqPost1}
	}
	\subfloat[][Step 3]{
		\fbox{\includegraphics[width=3.5cm]{pictures/PostExp_3.pdf}}
	    \label{fig:exseqPost2}
	}
	\subfloat[][Step 4 (\acro{prag} false)]{ 
		\fbox{\includegraphics[width=3.5cm]{pictures/PostExp_4.pdf}}
	    \label{fig:exseqPost3}
	}
	\caption{Example sequence for critical trial in
          post-experiment.}
	\label{fig:Trial-PostExp}
\end{figure}

@@

Backed up by our analyses, we concluded that both traditionalism and grammaticalism do not
predict our data under the set of auxiliary assumptions that we provided them with in
Section~\ref{sec:theories-predictions}. Traditionalism fails, as a core theory, to account for
high numbers of responses indicating local readings even for prosodically unmarked
\es-sentences. Grammaticalism as a core theory is trivially compatible with our data, because
it would be compatible with any observed preference order. But insofar as it builds on
disambiguation by logical strength, grammaticalism fails to predict the absence of answers
indicative of global readings and the relative abundance of answers indicative of literal
readings.

The clearest upshot of theoretical importance of this study is that strength-based meaning
selection does not seem to be a good disambiguation criterion for grammaticalism
\citep[c.f.][]{ChemlaSpector2010:Experimental-Ev}. This point is also relevant in general,
because disambiguation based on logical strength has been suggested in other domains as well
\citep[e.g.][]{DalrympleKanazawa1998:Reciprocal-Expr,Winter2001:Plural-Predicat,CobrerosEgre2012:Tolerant-Classi}.
We tentatively suggested an alternative selection principle for grammaticalism in terms of
``parsing economy,'' but it remains to be seen whether this is empirically successful in other
domains as well.

@@

We have argued that auditory presentation of sentence material is crucial if prosodic
information is hypothesized to influence availability and preferences of readings. We also
argued in passing that theoretical positions must be formulated in such a way that they make
testable predictions about reading preferences where different potential readings are
allowed. We should then try to obtain information about actual preference patterns by explicit
comparison with constructions whose preference structures are well-studied, ideally with some
model of the data generating process for optimal data analysis.
